Maenpuen et al.~Appendix S1-7: Sources and consequences of mismatch between leaf disc and whole-leaf leaf mass per area (LMA)

### 2021-12-15 ms/LMA_method_old.tex
first commit

Phisamai Maenpuen\textsuperscript{1,2,3}, Masatoshi
Katabuchi\textsuperscript{1,*}, Yusuke Onoda\textsuperscript{4}, Cong
Zhou\textsuperscript{1,2}, Jiao-Lin Zhang\textsuperscript{1,3*}, Ya-Jun
Chen\textsuperscript{1,3,5}

@@

\textsuperscript{5} Savanna Ecosystem Research Station, Xishuangbanna
Tropical Botanical Garden, Chinese Academy of Sciences, Yuanjiang,
Yunnan 6663300, China

@@

\textbf{METHODS:} We determined LMA using leaf discs and whole-leaf
(including petiole) for 334 woody species from a wide range of biomes
(tropics, subtropics, savanna, and temperate) to examine what extent
whole-leaf and leaf disc LMA match, whether the differences in the
estimates are associated with leaf size and thickness, and whether
intraspecifc variation for each species match.

@@

\textbf{CONCLUSIONS:} Mean values of leaf disc LMA are generally good
proxies for mean values of whole-leaf LMA with appropriate calibration
(9.4\% differences in our samples), but their accuracy depends on leaf
size, leaf thickness, and sizes of leaf punch. Therefore, using an
appropriate size of leaf punch is important for obtaining stable
estimates of leaf disc LMA that matches well with whole-leaf LMA.
Quantifying trait variation using leaf disc should be considered with
caution.

\textbf{KEY WORDS:} intraspecific variation, leaf density, leaf economic
spectrum, leaf heterogeneity, leaf punch, leaf size, leaf thickness,
petiole, specific leaf area, within-leaf variation

@@

LMA values can be determined by either 1) a whole leaf including or
excluding petioles, and 2) a leaf disc excluding all major veins or
petioles
(\protect\hyperlink{ref-Perez-Harguindeguy2013}{Pérez-Harguindeguy et
al., 2013}; \protect\hyperlink{ref-Kattge2020}{Kattge et al., 2020}).
Although some studies do not make clear what protocol was followed,
these differences in the measurement methods may lead to large
discrepancies in the estimates of LMA values because the major vein
allocation (major vein, which includes first to second or third-order
veins, volume per area) has been reported to be one of the main
determinants for the variation in LMA
(\protect\hyperlink{ref-John2017}{John et al., 2017}). Since larger
leaves tend to invest more of their mass into dense midribs and petiole
for support (\protect\hyperlink{ref-Niinemets2006}{Niinemets et al.,
2006}, \protect\hyperlink{ref-Niinemets2007}{2007}), and thinner leaves
have clearly visible and large-diameter veins with less uniform leaf
structure (i.e., kite-type leaves;
(\protect\hyperlink{ref-Grubb1986}{Grubb, 1986})), discrepancies in the
estimates of whole-leaf LMA and leaf disc LMA might be greater for
larger and thinner leaves. If leaf disc LMA largely and consistently
underestimates whole-leaf LMA, some calibrations may be required to
combine or compare those different estimates
(\protect\hyperlink{ref-Kraft2008}{Kraft et al., 2008};
\protect\hyperlink{ref-Onoda2011}{Onoda et al., 2011}). If divergencies
between the estimates based on the different methods are large and
inconsistent (i.e., low \emph{R\textsuperscript{2}} values), this is
difficult to calibrate and should inflate errors in the subsequent
analyses. To date, however, only a database from a single region, Panama
Plant Traits Database (\protect\hyperlink{ref-Wright2010}{Wright et al.,
2010}), is available in the TRY
(\protect\hyperlink{ref-Kattge2020}{Kattge et al., 2020}) that have both
estimates of LMA from small leaf discs and whole leaves including
petioles, limiting of our understanding of under what conditions
estimates of LMA from leaf discs are reliable estimates of whole-leaf
LMA.

@@

In this study, we aimed to investigate what extent whole-leaf (including
petiole) and leaf disc LMA match, whether the relationships between
whole-leaf LMA and leaf disc LMA varied in leaf sizes and leaf
thicknesses, and whether amounts of intraspecifc variations for each
match between whole-leaf and leaf disc based estimates. We collected a
total of 334 woody species from four biomes (tropics, subtropics,
savanna, and warm-temperate) to cover the wide range of geography. We
predicted that the overall relationship between whole-leaf LMA and disc
LMA would be weaker than previous reported LMA data
(\emph{R\textsuperscript{2}} = 0.92 for whole-leaves in Kraft et al.
(\protect\hyperlink{ref-Kraft2008}{2008}), \emph{R\textsuperscript{2}} =
0.92 for leaf laminas in Onoda et al.
(\protect\hyperlink{ref-Onoda2011}{2011})) because we collected data
with various leaf sizes from a wide range of climates. We also predicted
that species with larger and/or thinner leaves would show weaker
correlations. Finally, we predicted that amounts of intraspecifc
variations for each species do not match well between whole-leaf and
leaf disc based estimates.

@@

We used newly compiled individual-level plant datasets form a wide
ranges of biomes in China and Japan (Table S1) to examine the
relationship between whole-leaf and leaf disc-based estimates of leaf
traits. First, the Yunnan dataset is from three forest plots in Yunnan
province, Southwest China, which includes a tropical rain forest (TRF),
a hot-dry savanna ecosystem (HDS), and a subtropical evergreen wet
forest (STF), with a total of 141 species and 846 individuals. Second,
the Yakushima data is from a warm-temperate forest in Yakushima island,
Japan, with a total of 193 species and 519 individuals. In total, we
used 334 woody species and 1365 individuals. The methodologies for trait
sampling and measurement are slightly different between the Yunnan and
the Yakushima dataset, which we describe bellow.

\emph{Measurements of leaf disc and whole-leaf LMA}

At the TRF site, we collected leaf samples from trees within reach of a
canopy crane (88 m tall with a 60 m long boom). In the HDS, and STF
sites, we used a 12-m long pruner to collect samples for most of the
target species. In case of tall individuals that out of the reach of
pruner, we used a rope climbing to reach the canopy, then sampled
branches with the long pruner. In Yakushima, we used 15-m poles for our
leaf sampling. At least six sun-exposed healthy leaves were sampled from
each of 3 to 6 individuals for each species. Trait values were averaged
at the species- and individual-level for the analyses. For
compound-leaved species, we referred a leaflet (the minimum
photosynthetic unit) as a single leaf in our study.

To determine what extent LMA values were affected by the different
methods, LMA were determined (i) LMA based on a whole leaf including
petioles and (ii) LMA based on leaf disc avoiding thick veins. Three
0.6-cm diameter discs were taken from the base, middle and tip of each
leaf using a hole punch in the Yunnan dataset. Two 1.0-cm diameter discs
were taken from each leaf were in the Yakushima dataset. Fresh leaf area
(LA; cm\textsuperscript{-2}) of leaf materials (whole leaves including
petioles and midribs) was measured using a scanner and ImageJ software
by the R package \emph{LeafArea}
(\protect\hyperlink{ref-Katabuchi2015}{Katabuchi, 2015}). Fresh leaf
thickness (LT; mm) was measured using a micrometer in three points at
the base, middle and tip of the leaf using a micrometer (Mitutoyo
293-240, Mitutoyo Corporation, Kawasaki, Japan) with a precision at
0.001 mm. Thickness of each leaf disc was also measured for the Yunnan
dataset but not for the Yakushima dataset, which enables us to estimate
whole-leaf and leaf disc leaf tissue density (LD; g
cm\textsuperscript{-3}) separately in the Yunnan dataset. Leaf dry mass
(whole leaves including petioles and midribs, and leaf discs) was
recorded after drying 80°C for over 48 h to constant weight using an
electronic scale (Mettler-Toledo MS204TS, Columbus, OH, U.S.A.) with a
precision at 0.0001 g. In the subset of Yakushima dataset, leaf lamina
and petiole dry mass were also determined separately (81 out of 193
species) to quantify leaf support costs. Dry mass for leaf discs were
recorded 2-3 discs together for each leaf. Based on those measurements
of whole-leaf and leaf disc samples, LMA (g m\textsuperscript{-2}) was
calculated as the ratio of leaf dry mass to fresh leaf area, and LD was
calculated as the ratio of LMA to LT. We do not present results of LD in
the main text, because differences between whole-leaf LD and leaf disc
LD only depends on the ratio between whole-leaf LMA and leaf disc LMA
and measurement errors of leaf thickness (Appendix S2).

@@

Bivariate trait relationships between disc-based and whole-leaf
log-transformed LMA were summarized with variance explained
(\emph{R\textsuperscript{2}}) based on Pearson correlations and with
standardized major axis (SMA) regressions
(\protect\hyperlink{ref-Warton2006}{Warton et al., 2006}) across all the
samples and different leaf area and thickness groups. Species and
individual groups were divided into four categories based on the medians
of leaf area and leaf thickness across all species and all individuals,
respectively: thin and small-leaved, thin and large-leaved, thick and
small-leaved, thick and large-leaved species (individuals). Median
values are as following: small-leaved species (LA \textless{} 29.5
cm\textsuperscript{2}), thin-leaved species (LT \textless{} 0.28 mm),
small-leaved individuals (LA \textless{} 34.7 cm\textsuperscript{2}),
and thin-leaved individuals (LT \textless{} 0.26 mm).

@@

Disc-based estimates of species mean LMA matched well whole-leaf
estimates (Table 1; \emph{R\textsuperscript{2}} = 0.87 for all species).
Species with thicker and smaller leaves had stronger correlations
between disc-based and whole-leaf estimates of LMA (Table 1, Fig. 1).
There was a negative correlation between species mean leaf area and leaf
thickness (\emph{r} = -0.29, \emph{P} \textless{} 0.05, \emph{n} = 334).
When we pooled all the species, there were significant differences in
intercepts for species mean LMA (Table 1), suggesting that the
whole-leaf estimates of LMA were 9.4\% greater than disc-based estimates
in the arithmetic scale. Large-leaved species also had significant
positive intercepts, suggesting that whole-leaf estimates of LMA were
22.1\% greater than disc-based estimates in the arithmetic scale (Table
1). Intercepts did not differ for small-leaved species (Table 1).

Disc-based estimates of individual mean LMA also showed positive
correlations with whole-leaf estimates but their
\emph{R\textsuperscript{2}} values were smaller than those for the
species means (Fig. 2). Similar to the species-level pattern,
individuals with thicker and smaller leaves had stronger correlations
(Fig. 2). The relationships of the whole-leaf LMA : disc leaf LMA ratio
with total dry mass for the leaf disc (we weighted 2-3 discs together),
with leaf thickness, and with leaf area were heteroscedastic (unequal
variance, \emph{P} \textless{} 0.001; Appendix S3). Samples with small
total dry mass, thin leaves, and larger leaves tended to show greater
variance (Fig. 2; Appendix S3). The total dry mass of leaf disc was
positively correlated with both leaf thickness and leaf punch size
(Appendix S4). Amounts of intraspecific variation (CV; coefficient of
variation) only weakly matched between whole-leaf and disc-based
estimates (Fig. 3; \emph{R\textsuperscript{2}} = 0.31).

@@

Disc-based traits explained interspecific variations in whole-leaf LMA
(\emph{R\textsuperscript{2}} = 0.87, \emph{n} = 334) quite well, and the
strength of their relationship was approximately same or slightly weaker
than previously reported (\emph{R\textsuperscript{2}} = 0.92, \emph{n} =
409, for whole-leaves in Kraft et al.
(\protect\hyperlink{ref-Kraft2008}{2008}), \emph{R\textsuperscript{2}} =
0.92, \emph{n} = 364 for leaf laminas in Onoda et al.
(\protect\hyperlink{ref-Onoda2011}{2011})). Standardized major axis
(SMA) regression analysis showed significantly greater intercepts when
all the species were pooled (9.4\% differences) or large-leaved species
were used (22.1\% differences). Together, these findings suggest that
disc-based estimates can be generally used as good proxies for their
whole-leaf estimates with appropriate calibration. However, the strength
of correlations highly depended on leaf area, thickness (Fig. 1) and
sizes of leaf punches (Appendix S3 and S5).

@@

Meanwhile, our results are seemingly inconsistent with a previous study
that used thin and soft leaves and found high
\emph{R\textsuperscript{2}} values (\emph{R\textsuperscript{2}} = 0.92;
Kraft et al. (\protect\hyperlink{ref-Kraft2008}{2008})). Kraft et al.
(\protect\hyperlink{ref-Kraft2008}{2008}) used sapling leaves under
closed canopy that are thinner than adult leaves
(\protect\hyperlink{ref-Kitajima2010}{Kitajima and Poorter, 2010};
\protect\hyperlink{ref-Oktavia2020}{Oktavia and Jin, 2020}). Shade
leaves tend to have thinner leaves than sun leaves because of fewer
layers of palisade mesophyll cells
(\protect\hyperlink{ref-Terashima2001}{Terashima et al., 2001};
\protect\hyperlink{ref-Onoda2008}{Onoda et al., 2008}). The relatively
low \emph{R\textsuperscript{2}} values for thin-leaved species and
individuals in our study compared to the previous study may suggest not
only the effect of distinct veins on variance in LMA
(\protect\hyperlink{ref-John2017}{John et al., 2017}) but also technical
difficulty in measuring tiny weight of discs in lower LMA leaves. We
found that estimates of whole-leaf and leaf disc LMA diverged when total
dry mass of leaf discs was small (Fig. 2; Appendices S3), and thus
thin-leaved species, which have smaller dry mass (Appendix S4), tended
to showed the low \emph{R\textsuperscript{2}} values especially when we
used a smaller leaf punch (diameter: 0.6 cm; Appendix S5).

Even though we used the same species, the same individuals and the same
trait (LMA) to estimate the extents of intraspecific trait variation
(ITV), the explained variance of the relationship between whole-leaf and
disc-based estimate was surprisingly small (Fig. 3;
\emph{R\textsuperscript{2}} = 0.31). This is probably because leaf disc
samples have more sources of ITV than whole-leaf samples (i.e.,
variation among leaf discs within the same leaves). Although there are
not many studies that quantified trait variation within a single leaf, a
previous study that used a giant leaved species \emph{Alocasia
macrorrhiza}, suggests that LMA varied within leaves due to reduced
water supply and demand from midrib to outer regions
(\protect\hyperlink{ref-Li2013a}{Li et al., 2013}). Because calculating
the mean values generally diminishes the effect of within-sample
variation, the relationship between disc-based estimates of individual
mean LMA and its whole-leaf estimates will be stronger than the
relationship between leaf disc CV and whole-leaf CV. Thus, leaf disc and
whole-leaf estimates of species means will show the strongest
relationship. This is consistent with the \emph{R\textsuperscript{2}}
values in our results (Figs. 1-3). The mismatch of ITV between
whole-leaf and disc-based estimate could be ignored in community-level
ITV studies (i.e., \protect\hyperlink{ref-Siefert2015}{Siefert et al.,
2015}), which quantify the total variance of ITV within communities.
However, it is likely to change patterns for species-level ITV studies
(i.e., \protect\hyperlink{ref-Albert2010a}{Albert et al., 2010}), which
quantify extents of ITV for each species, since species with high ITV
based on leaf-disc estimates do not always have to show high ITV for
whole-leaf estimates. This problem will be more obvious when one uses
smaller leaf punches (Appendix S7).

The present study was designed to examine the extent to which whole leaf
and leaf disc LMA matched. Our study shows that mean values of leaf disc
LMA is generally a good proxy for mean values of whole-leaf LMA.
However, larger and thinner leaves tended to show the large
discrepancies between the leaf disc and whole-leaf estimates, and thus
when one needs to mix leaf disc LMA and whole-leaf LMA in a single
analysis (e.g., meta-analysis), appropriate calibrations are necessary.
Although our study was not designed to evaluate the effects of diameter
sizes of leaf punches, the sizes of leaf punches are also the important
sources of variance between whole-leaf and leaf disc LMA. This is
probably because small dry mass inflates measurement errors and small
leaf disc size inflates trait variation within leaves. Additionally,
although the mean values of leaf disc and whole-leaf LMA matched well,
quantifying trait variation using leaf discs should be considered with
caution, because variation among leaf discs within the same leaves
(including measurement errors) seems to be large. Leaf discs, which can
target leaf lamina avoiding leaf veins, are useful and essential for
biochemical determinations such as measurements of Rubisco activity
(\protect\hyperlink{ref-Parry2002}{Parry et al., 2002}), because the
photosynthetic activity occurs most in lamina rather than in veins for
C\textsubscript{3} plants (\protect\hyperlink{ref-Hibberd2002}{Hibberd
and Quick, 2002}; \protect\hyperlink{ref-Gao2018}{Gao et al., 2018}).
Leaf disc LMA is also the preferred method for quantifying leaf lamina
density, which avoid effects of veins. However, if the objective is to
use leaf disc LMA as an alternative for whole-leaf LMA, using an
appropriate size of leaf punch is important for obtaining stable
estimates of leaf disc LMA that matches well with whole-leaf LMA.

@@

The authors thank Lian-Bin Tao and Peng-Yun Yan for field work. This
work was funded by the National Natural Science Foundation of China
(41861144016, 32071735, 31570406, 32171507), the CAS `Light of West
China' Program, Yunnan Provincial Science and Technology Department
(2018HB068). PM was supported by financial support from CAS-TWAS
President's Fellowship for International Doctoral Students. MK was
supported by CAS President's International Fellowship Initiative
(2020FYB0003).

@@

\hypertarget{data-availability-statement}{%
\section{DATA AVAILABILITY
STATEMENT}\label{data-availability-statement}}

@@

\textbf{APPENDIX S1.} Site information.

\textbf{APPENDIX S2.} Generalization for a relationship between ratios
of disc-based and whole-leaf estimates of leaf mass per area (LMA), leaf
tissue density (LD), and leaf thickness (LT).

\textbf{APPENDIX S3.} Relationships between whole-leaf LMA : disc leaf
LMA and (a) total dry mass for the leaf disc, (b) leaf thickness, and
(c) leaf area.

\textbf{APPENDIX S4.} Summary of multiple regression for log-transformed
total dry mass of leaf discs.

\textbf{APPENDIX S5.} Relationships between species mean leaf mass per
area (LMA) determined by using whole leaves and leaf discs.

\textbf{APPENDIX S6.} Relationships between petiole to leaf dry mass
ratio and (a) leaf area and (b) whole-leaf to leaf disc LMA ratio.

\textbf{APPENDIX S7.} Relationships between coefficient of variation
(CV) in LMA determined by using whole leaves and leaf discs: (a) 0.6 cm
diameter leaf punch, and (b) 1.0 cm diameter leaf punch.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Grubb1986}{}}%
Grubb, P. 1986. Sclerophylls, pachyphylls, and pycnophylls: {The} nature
and significance of hard leaf surface. \emph{Insects and the Plant
Surface}: 137--150.

>>>>

Phisamai Maenpuen\textsuperscript{1,2,3}, Masatoshi
Katabuchi\textsuperscript{1,*}, Yusuke Onoda\textsuperscript{4}, Cong
Zhou\textsuperscript{1,2}, Jiao-Lin Zhang\textsuperscript{1,3*}, Ya-Jun
Chen\textsuperscript{1,3}

@@

@@

\textbf{METHODS:} We determined LMA using leaf discs and whole-leaf
(including petiole) for 334 woody species from across biomes (tropics,
subtropics, savanna, and temperate) to examine what extent whole-leaf
and leaf disc LMA match, whether the differences in the estimates are
associated with leaf size and thickness, and whether intraspecifc
variation for each species match.

@@

\textbf{CONCLUSIONS:} Mean values of leaf disc LMA are generally good
proxies for mean values of whole-leaf LMA with appropriate calibration
(9.4\% differences in our samples), but their accuracy depends on leaf
size, leaf thickness, and sizes of leaf punch. Therefore, using
appropriate sizes of leaf punch is important for obtaining stable
estimates of leaf disc LMA that matches well with whole-leaf LMA.
Quantifying trait variation using leaf disc should be considered with
caution.

\textbf{Key words:}

intraspecific variation, leaf density, leaf economic spectrum, leaf
heterogeneity, leaf punch, leaf size, leaf thickness, petiole, specific
leaf area, within-leaf variation

@@

LMA values are measured from either 1) a whole leaf including or
excluding petioles, and 2) a leaf disc excluding all major veins or
petioles
(\protect\hyperlink{ref-Perez-Harguindeguy2013}{Pérez-Harguindeguy et
al., 2013}; \protect\hyperlink{ref-Kattge2020}{Kattge et al., 2020}).
Although some studies do not make clear what protocol was followed,
these differences in the measurement methods may lead to large
discrepancies in the estimates of LMA values because the major vein
allocation (major vein, which includes first to second or third-order
veins, volume per area) has been reported to be one of the main
determinants for the variation in LMA
(\protect\hyperlink{ref-John2017}{John et al., 2017}). Since larger
leaves should invest more of their mass into dense midribs and petiole
for support (\protect\hyperlink{ref-Niinemets2006}{Niinemets et al.,
2006}, \protect\hyperlink{ref-Niinemets2007}{2007}), and thinner leaves
have clearly visible and large-diameter veins with less uniform leaf
structure {[}i.e., kite-type leaves; Grubb
(\protect\hyperlink{ref-Grubb1986}{1986}){]}, discrepancies in the
estimates of whole-leaf LMA and leaf disc LMA might be greater for
larger and thinner leaves. If leaf disc LMA largely underestimates
whole-leaf LMA, some calibrations may be required to combine or compare
those different estimates (\protect\hyperlink{ref-Kraft2008}{Kraft et
al., 2008}; \protect\hyperlink{ref-Onoda2011}{Onoda et al., 2011}). If
divergencies between the estimates based on the different methods are
large (i.e., low \emph{R\textsuperscript{2}} values), this is difficult
to calibrate and should inflate errors in the subsequent analyses. To
date, however, only a database from a single region, Panama Plant Traits
Database (\protect\hyperlink{ref-Wright2010}{Wright et al., 2010}), is
available in the TRY (\protect\hyperlink{ref-Kattge2020}{Kattge et al.,
2020}) that have both estimates of LMA from small leaf discs and whole
leaves including petioles, limiting of our understanding of under what
conditions estimates of LMA from leaf discs are reliable estimates of
whole-leaf LMA.

@@

In this study, we aimed to investigate to which extent whole-leaf
(including petiole) and leaf disc LMA match, whether the relationships
between whole-leaf LMA and leaf disc LMA varied in leaf sizes and leaf
thicknesses, and whether amounts of intraspecifc variations for each
match between whole-leaf and leaf disc based estimates. We collected a
total of 334 woody species from across biomes (tropics, subtropics,
savanna, and warm-temperate) to cover the wide range of geography. We
predicted that the overall relationship between whole-leaf LMA and disc
LMA would be weaker than previous reported LMA data
(\emph{R\textsuperscript{2}} = 0.92 for whole-leaves in Kraft et al.
(\protect\hyperlink{ref-Kraft2008}{2008}), \emph{R\textsuperscript{2}} =
0.92 for leaf laminas in Onoda et al.
(\protect\hyperlink{ref-Onoda2011}{2011})) because we collected data
with various leaf sizes from the wide range of climates. We also
predicted that species with larger and/or thinner leaves would show
weaker correlations. Finally, we predicted that amounts of intraspecifc
variations for each species do not match well between whole-leaf and
leaf disc based estimates.

@@

We used newly compiled individual-level plant datasets from broad
regions to examine the relationship between leaf traits estimates based
on whole-leaf and leaf disc across wide ranges of geography in China and
Japan (Table S1). First, the Yunnan dataset is from forest plots in
Yunnan province, Southwest China, which includes a tropical rain forest
(TRF), a hot-dry savanna ecosystem (HDS), and a subtropical evergreen
wet forest (STF), with a total of 141 species and 846 individuals.
Second, the Yakushima data is from a warm-temperate forest in Yakushima
island, Japan, with a total of 193 species and 519 individuals. In
total, we used 334 woody species and 1365 individuals. The trait
sampling methodology are slightly different between the Yunnan and the
Yakushima dataset, which we describe bellow.

\emph{Measurements of disc-based and whole-leaf LMA}

At the TRF site, we collected leaf samples from trees within reach of a
canopy crane (88 m tall with a 60 m long boom). In the HDS, and STF
sites, we used 12-m long pruner to collect samples for most of the
target species. In case of tall individuals that out of the reach of
pruner, we used a rope climbing to reach the canopy, then sampled
branches with the long pruner. In Yakushima, we used 15-m poles for our
leaf sampling. At least twelve sun-exposed healthy leaves were sampled
from 3 to 6 individuals for each species. Trait values were averaged at
the species- and individual-level for the analyses. For compound-leaved
species, we referred a leaflet (the minimum photosynthetic unit) as a
single leaf in our study.

To determine what extent LMA values were affected by the different
methods, LMA were determined (i) LMA based on a whole leaf including
petioles and (ii) LMA based on leaf disc avoiding thick veins. Three
small discs were taken from the base, middle and tip of each leaf using
a 0.6-cm diameter hole punch in the Yunnan dataset. Two large discs from
each leaf were taken using a 1.0-cm diameter hole punch in the Yakushima
dataset. Fresh leaf area (LA; cm\textsuperscript{-2}) of leaf materials
(whole leaves including petioles and midribs) was measured using scanner
and ImageJ software by the R package \emph{LeafArea}
(\protect\hyperlink{ref-Katabuchi2015}{Katabuchi, 2015}). Fresh leaf
thickness (LT; mm) was measured using a micrometer in three points at
the base, middle and tip of the leaf using a micrometer (Mitutoyo
293-240, Mitutoyo Corporation, Kawasaki, Japan) with a precision at
0.001 mm. Thickness of each leaf disc was also measured for the Yunnan
dataset but not for the Yakushima dataset, which enables us to estimate
whole-leaf and leaf disc leaf tissue density (LD; g
cm\textsuperscript{-3}) separately in the Yunnan dataset. Leaf dry mass
(whole leaves including petioles and midribs, and leaf discs) was
recorded after drying 80°C for over 48 h to constant weight using an
electronic scale (Mettler-Toledo MS204TS, Columbus, OH, U.S.A.) with a
precision at 0.0001 g. In the subset of Yakushima dataset, leaf lamina
and petiole dry mass were also determined separately (81 out of 193
species) to quantify leaf support costs. Dry mass for leaf discs were
recorded 2-3 discs together for each leaf. Based on those measurements
of whole-leaf and leaf disc samples, LMA (g m\textsuperscript{-2}) was
calculated as the ratio of leaf dry mass to fresh leaf area, and leaf
tissue density (LD; g cm\textsuperscript{-3}) was calculated as the
ratio of LMA to LT. We do not present results of LD in the main text,
because differences between whole-leaf LD and leaf disc LD only depends
on the ratio between whole-leaf LMA and leaf disc LMA and measurement
errors of leaf thickness (Appendix S2).

@@

Bivariate trait relationships between disc-based and whole-leaf
log-transformed LMA were summarized with variance explained
(\emph{R\textsuperscript{2}}) based on Pearson correlations and with
standardized major axis (SMA) regressions
(\protect\hyperlink{ref-Warton2006}{Warton et al., 2006}) across all the
samples and different leaf area and thickness groups. Species and
individual groups were divided into four categories based on the medians
of leaf area and leaf thickness across all species and all individuals,
respectively: thin and small-leaved, thin and large-leaved, thick and
small-leaved, thick and large-leaved species (individuals). Median
values are as following: small-leaved species (LA \textless{} 29.3
cm\textsuperscript{2}), thin-leaved species (LT \textless{} 0.28 mm),
small-leaved individuals (LA \textless{} 34.7 cm\textsuperscript{2}),
and thin-leaved individuals (LT \textless{} 0.26 mm).

@@

Disc-based estimates of species mean LMA matched well whole-leaf
estimates (Table 1; \emph{R\textsuperscript{2}} = 0.87 for all species).
Species with thicker and smaller leaves had stronger correlations
between disc-based and whole-leaf estimates of LMA (Table 1, Fig. 1).
There was a negative correlation between species mean leaf area and leaf
thickness (\emph{r} = -0.3, \emph{P} \textless{} 0.05, \emph{n} = 335).
When we pooled all the species, there were significant differences in
intercepts for species mean LMA (Table 1), suggesting that the
whole-leaf estimates of LMA were 9.4\% greater than disc-based estimates
in the arithmetic scale. Large-leaved species also had significant
positive intercepts, suggesting that whole-leaf estimates of LMA were
22.1\% greater than disc-based estimates in the arithmetic scale (Table
1). While intercepts did not differ for small-leaved species (Table 1).

Disc-based estimates of individual mean LMA also showed positive
correlations with whole-leaf estimates but their
\emph{R\textsuperscript{2}} values are smaller than those for the
species means (Fig. 2). Similar to the species-level pattern,
individuals with thicker and smaller leaves had stronger correlations
(Fig. 2). The relationships between whole-leaf LMA : disc leaf LMA and
total dry mass for the leaf disc (we weighted 2-3 discs together), leaf
thickness, and leaf area were heteroscedastic (unequal variance,
\emph{P} \textless{} 0.001). Samples with small total dry mass, thin
leaves, and larger leaves tended to show greater variance (Fig. 2;
Appendix S3). The total dry mass of leaf disc was positively correlated
with both leaf thickness and leaf punch size (Appendix S4). Amounts of
intraspecific variation (CV; coefficient of variation) only weakly
matched between whole-leaf and disc-based estimate (Fig. 3;
\emph{R\textsuperscript{2}} = 0.31).

@@

Disc-based traits explained interspecific variations in whole-leaf LMA
(\emph{R\textsuperscript{2}} = 0.87, \emph{n} = 335) quite well, and the
strength of their relationship was approximately same or slightly weaker
than previously reported (\emph{R\textsuperscript{2}} = 0.92, \emph{n} =
409, for whole-leaves in Kraft et al.
(\protect\hyperlink{ref-Kraft2008}{2008}), \emph{R\textsuperscript{2}} =
0.92, \emph{n} = 364 for leaf laminas in Onoda et al.
(\protect\hyperlink{ref-Onoda2011}{2011})). Standardized major axis
(SMA) regression analysis showed significantly greater intercepts when
all the species were pooled (9.4\% differences) or large-leaved species
were used (22.1\% differences). Together, these findings suggest that
disc-based estimates can be generally used as good proxies for their
whole-leaf estimates with appropriate calibration. However, the strength
of correlations highly depended on leaf area, thickness (Fig. 1) and
sizes of leaf punches (Appendix S3 and S5).

@@

Meanwhile, our results are seemingly inconsistent with a previous study
that used thin and soft leaves and found high
\emph{R\textsuperscript{2}} values (\emph{R\textsuperscript{2}} = 0.92;
Kraft et al. (\protect\hyperlink{ref-Kraft2008}{2008})). Kraft et al.
(\protect\hyperlink{ref-Kraft2008}{2008}) used sapling leaves under
closed canopy that are thinner than adult leaves
(\protect\hyperlink{ref-Kitajima2010}{Kitajima and Poorter, 2010};
\protect\hyperlink{ref-Oktavia2020}{Oktavia and Jin, 2020}). Shade
leaves tend to have thinner leaves than sun leaves because of the small
size and number of palisade mesophyll cells
(\protect\hyperlink{ref-Terashima2001}{Terashima et al., 2001};
\protect\hyperlink{ref-Onoda2008}{Onoda et al., 2008}). The relatively
low \emph{R\textsuperscript{2}} values for thin-leaved species and
individuals in our study compared to the previous study may suggest not
only the effect of distinct veins on variance in LMA
(\protect\hyperlink{ref-John2017}{John et al., 2017}) but also
measurement errors for those samples. We found that estimates of
whole-leaf and leaf disc LMA diverged when total dry mass of leaf discs
were small (Fig. 2; Appendices S3), and thus thin-leaved species, which
have smaller dry mass (Appendix S4), tended to showed the low
\emph{R\textsuperscript{2}} values especially when we used a smaller
leaf punch (diameter: 0.6 cm; Appendix S5).

Even though we used the same species, the same individuals and the same
trait (LMA) to estimate the extents of intraspecific trait variation
(ITV), the explained variance of the relationship between whole-leaf and
disc-based estimate was surprisingly small (Fig. 3;
\emph{R\textsuperscript{2}} = 0.31). This is probably because leaf disc
samples have more sources of ITV than whole-leaf samples (i.e.,
variation among leaf discs within the same leaves). Although there are
not many studies that quantified trait variation within a single leaf, a
previous study that used a giant leaved species \emph{Alocasia
macrorrhiza}, suggests that LMA varied within leaves due to reduced
water supply and demand from midrib to outer regions
(\protect\hyperlink{ref-Li2013a}{Li et al., 2013}). Because calculating
the mean values generally diminishes the effect of within-sample
variation, the relationship between disc-based estimates of individual
mean LMA and its whole-leaf estimates will be stronger than the
relationship between leaf disc CV and whole-leaf CV. Thus, leaf disc and
whole-leaf estimates of species means will show the strongest
relationship. This is consistent with the \emph{R\textsuperscript{2}}
values in our results (Figs. 1-3). The mismatch of ITV between
whole-leaf and disc-based estimate could be ignored for community-level
ITV studies (i.e., \protect\hyperlink{ref-Siefert2015}{Siefert et al.,
2015}), which quantify the total variance of ITV within communities.
However, it is likely to change patterns for species-level ITV studies
(i.e., \protect\hyperlink{ref-Albert2010a}{Albert et al., 2010}), which
quantify extents of ITV for each species, since species with high ITV
based on leaf-disc estimates do not always have to have high ITV for
whole-leaf estimates. This problem will be more obvious when one uses
smaller leaf punches (Appendix S7).

The present study was designed to examine the extent to which whole leaf
and leaf disc LMA matched. Our study shows that mean values of leaf disc
LMA is generally a good proxy for mean values of whole-leaf LMA.
However, larger and thinner leaves tended to show the large
discrepancies between the leaf disc and whole-leaf estimates, and thus
when one needs to mix leaf disc LMA and whole-leaf LMA in a single
analysis (e.g., meta-analysis), appropriate calibrations are necessary.
Although our study was not designed to evaluate the effects of diameter
sizes of leaf punches, the sizes of leaf punches are also the important
sources of variance between whole-leaf and leaf disc LMA. This is
probably because small dry mass inflates measurement errors and small
leaf disc size inflates trait variation within leaves. Additionally,
although the mean values of leaf disc and whole-leaf LMA matched well,
quantifying trait variation using leaf disc should be considered with
caution, because variation among leaf discs within the same leaves
(including measurement errors) seems to be large. Leaf discs, which can
target leaf lamina avoiding leaf veins, are useful and essential for
biochemical determinations such as measurements of Rubisco activity
(\protect\hyperlink{ref-Parry2002}{Parry et al., 2002}), because
photosynthetic is less active in veins than in mesophyll cells for
C\textsubscript{3} plants (\protect\hyperlink{ref-Hibberd2002}{Hibberd
and Quick, 2002}; \protect\hyperlink{ref-Gao2018}{Gao et al., 2018}).
Leaf disc LMA is also the preferred method for quantifying leaf lamina
density avoiding veins. However, if the objective is to use leaf disc
LMA as an alternative for whole-leaf LMA, using an appropriate size of
leaf punch is important for obtaining stable estimates of leaf disc LMA
that matches well with whole-leaf LMA.

@@

The authors thank Lian-Bin Tao and Peng-Yun Yan for field work. This
work was funded by the National Natural Science Foundation of China
(41861144016, 32071735, 31570406), the `Light of West China' Program of
CAS, Yunnan Provincial Science and Technology Department (2018HB068). PM
was supported by financial support from CAS-TWAS President's Fellowship
for International Doctoral Students. MK was supported by CAS President's
International Fellowship Initiative (2020FYB0003).

@@

\hypertarget{data-availability}{%
\section{DATA AVAILABILITY}\label{data-availability}}

@@

\textbf{APPENDIX S1-7.} Site information, generalization for a
relationship between ratios of disc-based and whole-leaf estimates of
leaf mass per area (LMA), leaf tissue density (LD), and complete figures
for mismatch between leaf disc and whole-leaf LMA.

@@

\leavevmode\vadjust pre{\hypertarget{ref-Grubb1986}{}}%
Grubb, P. 1986. Sclerophylls, pachyphylls, and pycnophylls: {The} nature
and significance of hard leaf surface. \emph{Insects and the plant
surface}: 137--150.
